---
title: "Gov_50 Final Project"
author: "Jonada Shukarasi"
description: "Public Opinion and US-Russia Relations"
output:
  distill::distill_article:
    self_contained: false
---
INTRODUCTION

The global landscape is marked by a myriad of socio-economic challenges, and the intersection of economic well-being, future outlook, and mental health outcomes remains an area for exploration. The central focus of this research project is to investigate the interplay between societal attitudes towards the future well-being of the next generation, economic satisfaction, and suicide rates on a global scale. The research question guiding this inquiry is: How do societal attitudes toward the future well-being of the next generation interact with economic satisfaction to influence suicide rates? The hypothesis posits that countries with higher economic satisfaction levels have higher optimistic attitudes about the financial prospects for the next generation, and in combination, are likely to exhibit lower suicide rates. Leveraging data obtained from surveys on economic satisfaction and future outlook, alongside global suicide rates compiled from official statistics, we seek to understand how societal attitudes towards the future well-being of the next generation interact with economic satisfaction to influence suicide rates. Suicide rates, the outcome variable of interest, will be measured through reliable global databases. The observed pattern supporting the hypothesis would reveal a positive correlation between economic satisfaction and positive attitudes toward the future, but a negative correlation between suicide rates and both economic satisfaction and positive attitudes toward the future. Conversely, a negative correlation between economic satisfaction and positive outlook would disprove the first part of the hypothesis, while evidence of lower suicide rates in countries with economic dissatisfaction and pessimistic outlooks would contradict the hypothesis, prompting a nuanced exploration of the intricate dynamics influencing societal mental health outcomes. The analysis will bw employing regression techniques to unravel the nuanced relationships between these variables.

```{r}
library(tidyverse)

economy <- read_csv('economy.csv')
future <- read_csv('future.csv')
suicide <- read_csv('suicide.csv')
```


DATA SECTION

In this study, we hypothesize a multifaceted relationship involving economic satisfaction, future outlook, and suicide rates on a global scale. First, we anticipate a positive correlation between economic satisfaction and optimistic attitudes toward the future, positing that countries with higher economic satisfaction levels are more likely to express positive expectations for the financial well-being of the next generation. Simultaneously, we expect that these optimistic attitudes, when coupled with higher economic satisfaction, will exhibit a negative correlation with suicide rates. The dual hypotheses aim to unravel the intricate interplay between economic factors, societal perceptions, and mental health outcomes on a global stage.

The unit of measurement for this research is the suicide rate, reflecting the prevalence of suicides per unit of population. The independent variables include economic satisfaction and future outlook, both derived from surveys conducted across diverse countries. The dependent variable, suicide rates, is sourced from global databases compiling official statistics. Using a regression analysis in R, we will explore the statistical relationships between these variables and test the hypotheses. Economic satisfaction and future outlook will be measured on ordinal scales derived from survey responses, capturing the nuanced opinions within each country. Suicide rates will be quantitatively measured, providing a robust indicator of mental health outcomes globally.

The dependent variable is the suicide rate, while economic satisfaction and future outlook serve as independent variables. Economic satisfaction is measured through survey responses, categorizing opinions on the economy as "good," "very good," "bad," or "very bad." Future outlook is similarly measured on an ordinal scale, with respondents indicating whether the future will be "better off," "the same," or "worse off." Suicide rates are quantitatively measured as the number of suicides per unit of population.

If we only consider the relationship between economic satisfaction and future outlook, the dependent variable would be "future outlook." In this context, economic satisfaction serves as an independent variable, and the focus is on how it influences or is associated with the perceptions and expectations regarding the future well-being of the next generation. The "future outlook" variable, measured on an ordinal scale as "better off," "the same," or "worse off," becomes the dependent variable that is influenced by the level of economic satisfaction.

We anticipate that countries with higher economic satisfaction levels and positive future outlooks will exhibit lower suicide rates, providing support for our hypotheses. Conversely, if observed patterns contradict our expectations, it will prompt a detailed exploration of the complex relationships between economic factors, societal attitudes, and mental health outcomes on a global scale. The study, conducted in R, aims to contribute valuable insights for understanding and addressing mental health challenges across diverse societies.



```{r}
# Filter Suicide Dataset so that we are only looking at the countries found in economy and future datasets
suicide <- suicide |>
  filter(country == 'Canada' |
         country == 'Belgium' |
         country == 'France' |
         country == 'Germany' |
         country == 'Greece' |
         country == 'Hungary' |
         country == 'Italy' |
         country == 'Netherlands (Kingdom of the)' |
         country == 'Poland' | 
         country == 'Spain' |
         country == 'Sweden' |
         country == 'United Kingdom of Great Britain and Northern Ireland' |
         country == 'Israel' |
         country == 'Australia' | 
         country == 'Japan' |
         country == 'Singapore' |
         country == 'Republic of Korea' |
         country == 'Malaysia')|>
  filter(population == 'Total' &
         age_range == 'Total years')
```

First, let us create plots to illustrate each country's economic satisfaction levels. 

```{r}
# Plot each opinion count for each country
very_good_plot <- economy |>
  ggplot(aes(x=country, y=very_good)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Economic Situation Opinion: Very Good', 
       x = 'Country',
       y = 'Count')

very_good_plot
```
```{r}
somewhat_good_plot <- economy |>
  ggplot(aes(x=country, y=somewhat_good)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Economic Situation Opinion: Somewhat Good', 
     x = 'Country',
     y = 'Count')

somewhat_good_plot
```

```{r}
somewhat_bad_plot <- economy |>
  ggplot(aes(x=country, y=somewhat_bad)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Economic Situation Opinion: Somewhat Bad', 
     x = 'Country',
     y = 'Count')

somewhat_bad_plot
```


```{r}
very_bad_plot <- economy |>
  ggplot(aes(x=country, y=very_bad)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Economic Situation Opinion: Very Bad', 
     x = 'Country',
     y = 'Count')

very_bad_plot
```

Now let us create plots that illustrate each country's future outlook. 

```{r}
better_off_plot <- future |>
  ggplot(aes(x=country, y=better_off)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Future Financial Opinion: Better Off', 
     x = 'Country',
     y = 'Count')

better_off_plot
```

```{r}
worse_off_plot <- future |>
  ggplot(aes(x=country, y=worse_off)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Future Financial Opinion: Worse Off', 
     x = 'Country',
     y = 'Count')

worse_off_plot
```

```{r}
same_plot <- future |>
  ggplot(aes(x=country, y=same)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Future Financial Opinion: Same', 
     x = 'Country',
     y = 'Count')

same_plot
```

And now we look at the suicide rate for every country. 

```{r}
suicide_rate_plot <- suicide |>
  ggplot(aes(x=country, y=suicide_rate)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8)) +
  labs(title = 'Suicide Rate by Country', 
     x = 'Country',
     y = 'Suicide Rate')

suicide_rate_plot
```



RESULTS

First need to mutate the three datasets to account for inconsistent country names between datasets. 

```{r}
# Merge Data
economy <- economy |>
  mutate(country_id = case_when(country == 'Canada' ~ 1, 
                                country == 'Belgium' ~ 2,
                                country == 'France' ~ 3,
                                country == 'Germany' ~ 4,
                                country == 'Greece' ~ 5,
                                country == 'Hungary' ~ 6,
                                country == 'Australia' ~ 7,
                                country == 'Japan' ~ 8,
                                country == 'Malaysia' ~ 9,
                                country == 'UK' ~ 10,
                                country == 'Netherlands' ~ 11,
                                country == 'South Korea' ~ 12,
                                country == 'Singapore' ~ 13,
                                country == 'Italy' ~ 14,
                                country == 'Spain' ~ 15,
                                country == 'Sweden' ~ 16,
                                country == 'Poland' ~ 17,
                                country == 'Israel' ~ 18))

future <- future |>
  mutate(country_id = case_when(country == 'Canada' ~ 1, 
                                country == 'Belgium' ~ 2,
                                country == 'France' ~ 3,
                                country == 'Germany' ~ 4,
                                country == 'Greece' ~ 5,
                                country == 'Hungary' ~ 6,
                                country == 'Australia' ~ 7,
                                country == 'Japan' ~ 8,
                                country == 'Malaysia' ~ 9,
                                country == 'UK' ~ 10,
                                country == 'Netherlands' ~ 11,
                                country == 'South Korea' ~ 12,
                                country == 'Singapore' ~ 13,
                                country == 'Italy' ~ 14,
                                country == 'Spain' ~ 15,
                                country == 'Sweden' ~ 16,
                                country == 'Poland' ~ 17,
                                country == 'Israel' ~ 18))

suicide <- suicide |>
  mutate(country_id = case_when(country == 'Canada' ~ 1, 
                                country == 'Belgium' ~ 2,
                                country == 'France' ~ 3,
                                country == 'Germany' ~ 4,
                                country == 'Greece' ~ 5,
                                country == 'Hungary' ~ 6,
                                country == 'Australia' ~ 7,
                                country == 'Japan' ~ 8,
                                country == 'Malaysia' ~ 9,
                                country == 'United Kingdom of Great Britain and Northern Ireland' ~ 10,
                                country == 'Netherlands (Kingdom of the)' ~ 11,
                                country == 'Republic of Korea' ~ 12,
                                country == 'Singapore' ~ 13,
                                country == 'Italy' ~ 14,
                                country == 'Spain' ~ 15,
                                country == 'Sweden' ~ 16,
                                country == 'Poland' ~ 17,
                                country == 'Israel' ~ 18))

merged_data <- economy |>
  inner_join(future, by = 'country_id') |>
  inner_join(suicide, by = 'country_id')

# I made a good_opinion and bad_opinion column which combines the very good and somewhat good responses and does the same for the bad ones
merged_data <- merged_data |>
  mutate(good_opinion = very_good + somewhat_good, 
         bad_opinion = very_bad + somewhat_bad)

view(merged_data)
```

Now let us look at the relationship between economic satisfaction and future outlook. First let's graph the relationship between "better off" outlook and "very good" economic satisfaction, then "better off" and "somewhat good", and finally general good opinion and general satisfaction. We will repeat the same thing for negative responses. 

```{r}
# Plots of Relationships Between Future Opinion and Economic Opinion

bo_vg_plot <- merged_data |>
  ggplot(aes(x = better_off, y = very_good)) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red')
  labs(title = 'Better Off Future Opinion vs Very Good Economic Opinion',
       x = 'Better Off Responses', 
       y = 'Very Good Responses')
  
bo_vg_plot


bo_sg_plot <- merged_data |>
  ggplot(aes(x = better_off, y = somewhat_good)) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = 'Better Off Future Opinion vs Somewhat Good Economic Opinion',
       x = 'Better Off Responses', 
       y = 'Somewhat Good Responses')
  
bo_sg_plot


bo_g_plot <- merged_data |>
  ggplot(aes(x = better_off, y = good_opinion)) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = 'Better Off Future Opinion vs Good Economic Opinion',
       x = 'Better Off Responses', 
       y = 'Good Responses')
  
bo_g_plot


wo_vb_plot <- merged_data |>
  ggplot(aes(x = worse_off, y = very_bad)) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = 'Worse Off Future Opinion vs Very Bad Economic Opinion',
       x = 'Worse Off Responses', 
       y = 'Very Bad Responses')
  
wo_vb_plot


wo_sb_plot <- merged_data |>
  ggplot(aes(x = worse_off, y = somewhat_bad)) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = 'Worse Off Future Opinion vs Somewhat Bad Economic Opinion',
       x = 'Worse Off Responses', 
       y = 'Somewhat Bad Responses')
  
wo_sb_plot


bo_b_plot <- merged_data |>
  ggplot(aes(x = better_off, y = bad_opinion)) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = 'Better Off Future Opinion vs Bad Economic Opinion',
       x = 'Better Off Responses', 
       y = 'Bad Responses')
  
bo_b_plot
```

Better Off Future Opinion vs. Economic Opinion Plots:

Better Off Future Opinion vs. Very Good Economic Opinion: The plot shows the relationship between respondents who believe the future will be better off and those who have a very good opinion about the current economic state. The positive slope of the regression line suggests that there is a tendency for respondents with a positive outlook on the future to also hold a very good opinion about the current economic situation.

Better Off Future Opinion vs. Somewhat Good Economic Opinion: This plot explores the association between a positive future outlook and a somewhat good opinion about the economy. The positive slope of the regression line indicates that individuals with an optimistic view of the future tend to also perceive the current economic situation as somewhat good.

Better Off Future Opinion vs. Good Economic Opinion: Similar to the previous plots, this one illustrates the correlation between a positive future outlook and a good opinion about the current economy. The positive slope suggests a relationship where those expecting a better future also tend to have a favorable opinion about the economy.


Worse Off Future Opinion vs. Economic Opinion Plots:

Worse Off Future Opinion vs. Very Bad Economic Opinion: Examining respondents who believe the future will be worse off against those with a very bad opinion about the economy, the positive slope suggests a connection between a negative future outlook and a very bad opinion about the current economic state.

Worse Off Future Opinion vs. Somewhat Bad Economic Opinion: This plot explores the link between a negative future outlook and a somewhat bad opinion about the current economy. The positive slope indicates that individuals expecting a worse future are more likely to view the current economic situation as somewhat bad.

Better Off Future Opinion vs. Bad Economic Opinion: The last plot in this category assesses the relationship between a negative future outlook and a bad opinion about the current economic situation. The positive slope suggests that individuals with a pessimistic view of the future are more likely to hold a negative opinion about the economy.

The plots reveal interesting patterns in the relationships between future outlook and economic opinions. Overall, there is a consistent positive trend where a positive future outlook aligns with more favorable opinions about the current economic state, and conversely, a negative future outlook corresponds with less favorable economic opinions. These findings suggest an interconnectedness between perceptions of the future and evaluations of the current economic climate.






















































